---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setoptions, echo=FALSE}
options(scipen = 1, digits = 2)
```


## Loading and preprocessing the data
```{r, results='hide'}
data <- read.csv(unz("activity.zip", "activity.csv"))
data$date <- as.Date(data$date, format = '%Y-%m-%d')
```


## What is mean total number of steps taken per day?
```{r}
stepsperday <- aggregate(steps ~ date, data = data, FUN = sum, na.rm = TRUE)
hist(stepsperday$steps)
```

Average (mean) number of steps taken per day
```{r}
mean(stepsperday$steps)
```

Median number of steps taken per day
```{r}
median(stepsperday$steps)
```


## What is the average daily activity pattern?
```{r}
stepsperinterval <- aggregate(steps ~ interval, data = data, FUN = mean, na.rm = TRUE)
with(stepsperinterval, plot(interval, steps, type = 'l'))
```

The 5-minute interval with most steps on average
```{r}
stepsperinterval[which.max(stepsperinterval$steps), ]$interval
```


## Imputing missing values
Number of rows with NAs
```{r}
sum(!complete.cases(data))
```

### Estimation strategy for missing values
For the `(date, interval)`s where `steps` is `NA`, I'm assigning the interval's average number of steps across all the dates.
```{r}
imputemissing <- function(val, idx) {
    if (is.na(val)) 
        stepsperinterval$steps[idx]
    else
        val
}
impdata <- data.frame(data)
impdata$steps <- mapply(imputemissing, val = impdata$steps, idx = as.numeric(rownames(impdata)))

impstepsperday <- aggregate(steps ~ date, data = impdata, FUN = sum)
hist(stepsperday$steps)
```

Average (mean) number of steps taken per day
```{r}
mean(impstepsperday$steps)
```

Median number of steps taken per day
```{r}
median(impstepsperday$steps)
```

Notice that imputed values do not seem to have much effect on the mean and median of the number of steps taken per day.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(ggplot2)

impdata$daytype <- factor(
    grepl("S(at|un)", weekdays(impdata$date, abbr = TRUE)), 
    levels = c("FALSE", "TRUE"), 
    labels = c("weekday", "weekend"),
)
aggdata <- aggregate(steps ~ interval * daytype, data = impdata, FUN = mean)

qplot(interval, steps, data = aggdata, facets = . ~ daytype, geom = c("line"))
```
